# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

```{r load and process data}

  activity <- read.csv("./data/activity.csv",header=TRUE,colClasses=c("numeric","Date","numeric"),na.strings="NA")

  library(lubridate)
  activity$date <- ymd(activity$date) ##transform column date as a date

```

## What is mean total number of steps taken per day?

```{r hist of mean total number of steps}

  # Sum of total steps per day

    totalsteps <- data.frame(totalsteps = sapply(split(activity$step,activity$date),sum,na.rm=TRUE))

  ## Histogram of the total number of   steps taken each day
  
    library(ggplot2)    

    qplot(totalsteps,data = totalsteps,xlab="Total steps per day",ylab="Count", 
          main="Histogram of the total number of   steps taken each day")

    meansteps <- mean(totalsteps$totalsteps)
    mediansteps <- median(totalsteps$totalsteps)
  

```

The mean total number of steps taken by day is **`r meansteps`**  
The median  total number of steps taken by day is **`r mediansteps`**

## What is the average daily activity pattern?

```{r average daily activitt}

  ## average steps across intervals

    averagesteps <- data.frame(average = sapply(split(activity$step,
                        activity$interval),mean,na.rm=TRUE))
    averagesteps$interval <- as.numeric(row.names(averagesteps))

  ## plot 

    ggplot(averagesteps,aes(interval,average)) + geom_line()

  ## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

    maxsteps <- subset(averagesteps$interval,
         averagesteps$average == max(averagesteps$average))
  

```

The interval with contains the maximum number of steps is the interval **`r maxsteps`**

## Imputing missing values

```{r number of missing value}

  missing <- sum(is.na(activity$steps))

```

The number of rows with missing value is **`r missing`**

```{r impute mean}

  # for each missing value in steps, the mean of the mean of the corresponding  5-minute interval is imputed

    activity2 <- activity
  
    for (i in 1:nrow(activity2))
      {
        if (is.na(activity2[i,1]) == TRUE)
          
          {
            x <- subset(activity2,activity2[,3] == activity2[i,3])
            activity2[i,1] <- mean(x$steps,na.rm=TRUE)
          }
      }

  # Sum of total steps per day

    totalsteps2 <- data.frame(totalsteps = 
                sapply(split(activity2$step,activity2$date),sum,na.rm=TRUE))

  ## Histogram of the total number of   steps taken each day
  
    library(ggplot2)    

    qplot(totalsteps,data = totalsteps2,xlab="Total steps per day",ylab="Count",
          main="Histogram of the total number of   steps taken each day")

    meansteps2 <- mean(totalsteps2$totalsteps)
    mediansteps2 <- median(totalsteps2$totalsteps)

```

**Dataset with missing values**  
The mean total number of steps taken by day is **`r meansteps`**  
The median  total number of steps taken by day is **`r mediansteps`**  
**Dataset without missing values**  
The mean total number of steps taken by day is **`r meansteps2`**  
The median  total number of steps taken by day is **`r mediansteps2`**  

## Are there differences in activity patterns between weekdays and weekends?

```{r}

  # create levels

    activity2$weekdays <- as.factor(weekdays(activity2$date,abbreviate=TRUE))
    levels(activity2$weekdays) <-c("weekend","weekday","weekday","weekday",
                       "weekday","weekend","weekday")

  #   create subsets for weekdays and weekend
  
    weekday.act <- data.frame(subset(activity2,activity2$weekdays == "weekday"))
    weekend.act <- data.frame(subset(activity2,activity2$weekdays == "weekend"))

  # compute means
  
    weekday.ave <- data.frame(average = sapply(split(weekday.act$step,
                   weekday.act$interval),mean,na.rm=TRUE))
    weekday.ave$interval <- as.numeric(row.names(weekday.ave))
    weekday.ave$weekdays <- as.factor(c("weekdays"))
    
    weekend.ave <- data.frame(average = sapply(split(weekend.act$step,
                   weekend.act$interval),mean,na.rm=TRUE))
    weekend.ave$interval <- as.numeric(row.names(weekend.ave))
    weekend.ave$weekdays <- as.factor(c("weekend"))
    
    average <- rbind(weekday.ave,weekend.ave)

    ggplot(average,aes(interval,average)) +facet_grid (weekdays~.) + geom_line()

```


